\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-06-23 paper.qmd
done with 2D

#### Computing hexagon grid configurations

The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and hexagon id, $h = 1, \dots, b$. Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$. The lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid. The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer ($q$) corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data. The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).

@@

Observations are grouped into bins based on their nearest centroid. This can be defined using the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, where
$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$, mapping observation $i$ into $H_h = \{i| u(i) = h\}$.

When $k = 2$, the hexagonal bin centroid is used (@fig-NLDR-scurve (c)). Any measure of center, such as a mean or weighted mean of the points within each hexagon, could be used to represent the hexagon.

@@

#### Indicating neighbors by line segments connecting centroids

@@

### Displaying the model in \pD{}

>>>>

#### Computing hexagon grid configuration

The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and a unique hexagon id, $h = 1, \dots, b$. Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.  (Note that, we detail the algorithm for constructing the grid and binning the data because we found that there was no readily available implementation that could produce what was needed for this project.)

To make the grid, a buffer parameter ($q$) is set as a proportion of the plot space. By default this is 0.1. A value should be chosen to provide at buffer of a full hexagon width ($a_1$) and height ($a_2$) for the grid to extend beyond the data. The lower left position where the grid starts is defined as $(s_1, s_2)$, corresponding to the centroid of the lowest left hexagon, which should be below the minimum data value. This is computed as one buffer unit, $q$ below the minimum data value, by default. The user provides the preferred number of bins in the horizontal direction, $b_1$, and the algorithm computes the appropriate number of bins in the vertical direction, $b_2$.

@@

Observations are grouped into bins based on their nearest centroid. This produces a reduction in size of the data from $n$ to $m$, where $m\leq b$ (total number of bins). This can be defined using the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, where
$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$, mapping observation $i$ into $H_h = \{i| u(i) = h\}$.

By default, the bin centroid is used for describing a hexagon (as done in @fig-NLDR-scurve (c)), but any measure of center, such as a mean or weighted mean of the points within each hexagon, could be used. The bin centers, and the binned data, are the two important components needed to render the model representation in high dimensions.

@@

#### Indicating neighborhood

@@

### Rendering the model in \pD{}

### 2024-06-23 paper.tex
done with 2D

\subsubsection{Computing hexagon grid
configurations}\label{computing-hexagon-grid-configurations}

The \gD{} hexagon grid is defined by the number of bins in each
direction \((b_1, b_2)\), giving total number of bins as
\(b = b_1 \times b_2\), and hexagon id, \(h = 1, \dots, b\). Each
hexagon, \(H_h\) is uniquely described by centroid,
\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). The lower left position where the
grid starts at \((s_1, s_2)\), which correspond to the lowest left
centroid. The values of \(s_i\) need to be below their respective
minimum variable values, and could be a full bin lower, to allow a
buffer (\(q\)) corresponding to a full hexagon width (\(a_1\)) and
height (\(a_2\)) around the data. The values of \(b_i\) are variables to
be computed that define the reduction in size of the data (\(n\) to
\(m\)).

@@

Observations are grouped into bins based on their nearest centroid. This
can be defined using the function
\(u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}\), where
\(u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}\),
mapping observation \(i\) into \(H_h = \{i| u(i) = h\}\).

When \(k = 2\), the hexagonal bin centroid is used
(Figure~\ref{fig-NLDR-scurve} (c)). Any measure of center, such as a
mean or weighted mean of the points within each hexagon, could be used
to represent the hexagon.

\subsubsection{Indicating neighbors by line segments connecting
centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}

@@

\subsection{\texorpdfstring{Displaying the model in
\pD{}}{Displaying the model in }}\label{displaying-the-model-in}

>>>>

\subsubsection{Computing hexagon grid
configuration}\label{computing-hexagon-grid-configuration}

The \gD{} hexagon grid is defined by the number of bins in each
direction \((b_1, b_2)\), giving total number of bins as
\(b = b_1 \times b_2\), and a unique hexagon id, \(h = 1, \dots, b\).
Each hexagon, \(H_h\) is uniquely described by centroid,
\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). (Note that, we detail the algorithm
for constructing the grid and binning the data because we found that
there was no readily available implementation that could produce what
was needed for this project.)

To make the grid, a buffer parameter (\(q\)) is set as a proportion of
the plot space. By default this is 0.1. A value should be chosen to
provide at buffer of a full hexagon width (\(a_1\)) and height (\(a_2\))
for the grid to extend beyond the data. The lower left position where
the grid starts is defined as \((s_1, s_2)\), corresponding to the
centroid of the lowest left hexagon, which should be below the minimum
data value. This is computed as one buffer unit, \(q\) below the minimum
data value, by default. The user provides the preferred number of bins
in the horizontal direction, \(b_1\), and the algorithm computes the
appropriate number of bins in the vertical direction, \(b_2\).

@@

Observations are grouped into bins based on their nearest centroid. This
produces a reduction in size of the data from \(n\) to \(m\), where
\(m\leq b\) (total number of bins). This can be defined using the
function
\(u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}\), where
\(u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}\),
mapping observation \(i\) into \(H_h = \{i| u(i) = h\}\).

By default, the bin centroid is used for describing a hexagon (as done
in Figure~\ref{fig-NLDR-scurve} (c)), but any measure of center, such as
a mean or weighted mean of the points within each hexagon, could be
used. The bin centers, and the binned data, are the two important
components needed to render the model representation in high dimensions.

\subsubsection{Indicating neighborhood}\label{indicating-neighborhood}

@@

\subsection{\texorpdfstring{Rendering the model in
\pD{}}{Rendering the model in }}\label{rendering-the-model-in}
